---
title: COVID-19 vaccine interest in Malaysia
author: tengku-hanis
date: '2021-10-17'
slug: covid-19-vaccine-interest-in-malaysia
categories:
  - R
  - text analysis
tags:
  - Vaccine
  - text analysis
subtitle: ''
summary: ''
authors: []
lastmod: '2021-10-17T12:36:00+08:00'
featured: no
image:
  caption: ''
  focal_point: ''
  preview_only: no
projects: []
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
```

We are going to do a basic google trends search using `gtrendsR` package and do some plotting with `ggplot2`.

These are the required packages.
```{r}
library(gtrendsR)
library(tidyverse)
```

Run `gtrends()` function to search our keywords of interest (i.e; type of vaccine). So far, we only used [4 type of vaccines](https://covidnow.moh.gov.my/vaccinations/) in Malaysia.
```{r}
vaccine <- gtrends(c("pfizer", "astrazeneca", "sinovac", "cansino"), geo = "MY")
```

Then, plot our keywords.
```{r}
plot(vaccine)
```
Probably, it's better if we filter our date to when the COVID-19 pandemic started, which is around March 2020.
```{r}
vaccine$interest_over_time %>% 
  group_by(keyword) %>% 
  filter(hits != "<1" & date > as.Date("2020-03-01")) %>% 
  mutate(hits = as.numeric(hits), 
         date = as.Date(date)) %>% 
  ggplot() + 
  geom_line(aes(x = date, y = hits, color = keyword), size = 0.8) +
  theme_minimal() +
  labs(title = "COVID-19 vaccine interest in Malaysia", y = "Search hits", x = "Date") +
  scale_x_date(date_breaks = "4 month")
```

So, AstraZeneca vaccine is of high interest, probably due to infamous blood clotting issue. Next, we can also get the search keywords based on the states.
```{r}
vaccine$interest_by_region %>% 
  group_by(location) %>% 
  ggplot(aes(location, hits, fill = keyword)) +
  geom_col(alpha = 0.8) +
  coord_flip() +
  theme_minimal() +
  scale_fill_viridis_d() +
  labs(title = "COVID-19 vaccine interest in Malaysia by states", y = "Search hits", x = "")
```

Lastly, we can plot the search keywords based on the city.
```{r}
vaccine$interest_by_city %>% 
  group_by(location) %>% 
  drop_na() %>% 
  ggplot(aes(location, hits, fill = keyword)) +
  geom_col(alpha = 0.8) +
  coord_flip() +
  theme_minimal() +
  scale_fill_viridis_d() +
  labs(title = "COVID-19 vaccine interest in Malaysia by cities", y = "Search hits", x = "")
```

`gtrendsR` with just a bit of plots certainly very useful if we want to gauge certain issues in the community. 
