---
title: Base R vs tidyverse
author: Tengku Hanis
date: '2021-05-04'
slug: []
categories:
  - R
tags:
  - tidyverse
  - base R
subtitle: ''
summary: ''
authors: []
lastmod: '2021-05-04T16:48:49+08:00'
featured: no
image:
  caption: ''
  focal_point: ''
  preview_only: no
projects: []
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

First of all, this write up is mean for a beginner in R. 

Things can be done in many ways in R. In facts, R has been very flexible in this regard compared to other statistical softwares. Basic things such as selecting a column, slicing a row, filtering a data based on certain condition can be done using a base R function. However, all these things can also be done using a tidyverse approach. 

[Tidyverse](https://www.tidyverse.org/) basically, a collection of packages that can be loaded in a line of function. 

```{r, eval=FALSE}
library(tidyverse)
```

Tidyverse is developed by "RStudio people" pioneered  by [Hadley Wickham](http://hadley.nz/), which means that these packages will be continuously maintained and updated. 

So, without further ado, these are the comparisons between these two approaches for some very basic thingy: 
1. Select or deselect a column and a row
```{r, eval=FALSE}
# Base R
iris[1:5, c("Sepal.Length", "Sepal.Width")]
iris[1:5,c(1,2)] # similar to above
iris[1:5, -1]

# Tidyverse
iris %>% 
  select(Sepal.Length, Sepal.Width) %>% 
  slice(1:5)
iris %>% 
  select(-Sepal.Length) %>% 
  slice(1:5)
```

2. Filter based on condition
```{r, eval=FALSE}
# Base R
iris[iris$Species == "setosa", ]

# Tidyverse
iris %>% 
  filter(Species == "setosa")
```

3. Mutate (transmute replace the variable)
```{r,  eval=FALSE}
# Base R
iris$SL_minus10 <- iris$Sepal.Length - 10

# Tidyverse
iris %>% 
  mutate(SL_minus10 = Sepal.Length - 10)
```

4. Sort variable
```{r,  eval=FALSE}
# Base R
iris[order(-iris$Sepal.Width),]

# Tidyverse
iris %>% 
  arrange(desc(Sepal.Length))
```
5. Group by (and get mean for variable Sepal.Width)
```{r, eval=FALSE}
# Not really base R
doBy::summaryBy(Sepal.Width~Species, iris, FUN = mean) 

# Tidyverse
iris %>% 
  group_by(Species) %>% 
  summarise(mean_SW = mean(Sepal.Width))
```

6. Rename variable
```{r, eval=FALSE}
# Base R
colnames(iris)[6] <- "hanis"

# Tidyverse
iris %>% 
  rename(Species = hanis)
```

So, that's it. Overall, tidyverse give a clarity in understanding the code as it reads from left to right. On the contrary, the base R approach reads from inside to outside, especially for a more complicated code.
