---
title: Data exploration in R
author: Tengku Hanis
date: '2021-08-22'
slug: data-exploration-in-r
categories:
  - R
tags:
  - Data exploration
subtitle: ''
summary: ''
authors: []
lastmod: '2021-08-22T05:10:54+08:00'
featured: no
image:
  caption: ''
  focal_point: ''
  preview_only: no
projects: []
---
```{r setup, eval=TRUE, echo=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
```

These are some of the packages that I find useful for data exploration. Basically, this post serves more as my note for future reference. I will list out packages (and some awesome functions from that particular package) rather than specific functions. Further, base R and tidyverse packages will not be included specifically in this list.

Load supporting packages

```{r, warning=FALSE, message=FALSE}
library(tidyverse)
```

The data we are going to use is from dlookr package:

```{r, echo=FALSE, warning=FALSE, message=FALSE}
library(dlookr)
```

```{r}
glimpse(heartfailure)
```
We will create a few NAs in our data.
```{r}
set.seed(2021)
heartfailure[sample(seq(nrow(heartfailure)), 20), "age"] <- NA
heartfailure[sample(seq(nrow(heartfailure)), 10), "sex"] <- NA
```

**1)  dataMaid**

```{r}
library(dataMaid)
```

One of the very useful function in dataMaid is makeDataReport which give report on the data. By default it will give a pdf output, but other output option such as word and html can be set.

```{r, eval=FALSE}
makeDataReport(heartfailure, replace = T)
```

This is the output exampple in [pdf](https://tengkuhanis.netlify.app/files/dataMaid_heartfailure.pdf).

**2)  DataExplorer**

```{r}
library(DataExplorer)
```

General visualization:

```{r}
heartfailure %>% plot_intro()
```

Since we have missing data, we can further visualize it:

```{r}
heartfailure %>% plot_missing()

heartfailure %>% profile_missing()
```

We can also do a correlation plot

```{r}
heartfailure %>% 
  select_if(is.numeric) %>% 
  drop_na() %>% 
  plot_correlation()
```

However, I do think correlation plot from corrplot packages gives a better and clean plot. Here is a plot from corrplot.

```{r}
library(corrplot)

heartfailure %>% 
  select_if(is.numeric) %>% 
  drop_na() %>% 
  cor() %>% 
  corrplot(type = "upper")
```

Finally, we can get an overall html report from DataExplorer package using the function `create_report()`.

**3) dlookr**

```{r}
library(dlookr)
```

We can assess normality of the data using this package. The code below will plot normality for all numeric variable.
```{r, eval=FALSE}
heartfailure %>% 
  plot_normality()
```
However, for the sake of the simplicity in this post, we will run only for one variable.
```{r}
heartfailure %>% 
  plot_normality(age)
```

We can also get a correlation matrix plot from this package, and no need to remove the NAs and filter the numeric variable before running the function.
```{r}
heartfailure %>% 
  plot_correlate()
```

Lastly, from dlookr we can get the overall report of the data exploration in pdf (and other formats as well). This report is quite comprehensive, have a [look](https://tengkuhanis.netlify.app/files/EDA_Paged_Report.pdf). 
```{r, eval=FALSE}
heartfailure %>% 
  eda_paged_report(target = "death_event")
```

**4) outliertree**

This package identify outlier using a decision tree. I will not go in detail about the approach, but for those who want to read [further](https://arxiv.org/abs/2001.00636).  
```{r}
library(outliertree)
outlier.tree(heartfailure)
```

We can further explore the detected outliers using histogram and boxplot. Let's do for variable creatinine.
```{r}
# histogram
hist(heartfailure$creatinine, breaks = 50, col = "navy",
     xlab = "Creatinine", 
     main = "Creatinine level")

# boxplot
boxplot(heartfailure$creatinine)
```

Probably in the future I will delve into more detail about outlier detection and any awesome packages in R related to it. If I ever written any post about it, I will link it here.

## Conclusion

These are some useful package that I find. I may edit this post in the future to add more additional data exploration package. Furthermore, there are shiny apps for data exploration as well, though I think it's better to sticks with coded approach in data analysis/exploration. Thus, I did not explore those apps in this post. Another thing to remember is to set the variable type accordingly prior to the data exploration.

Hope this is useful!

References:\
<https://github.com/ekstroem/dataMaid>\
<https://finnstats.com/index.php/2021/05/04/exploratory-data-analysis/>\
<https://cran.r-project.org/web/packages/dlookr/vignettes/EDA.html>\
<https://cran.r-project.org/web/packages/outliertree/vignettes/Introducing_OutlierTree.html>
