---
title: Fitted vs predict in R
author: ''
date: '2022-01-09'
slug: fitted-vs-predict-in-r
categories:
  - applied statistics
  - R
tags:
  - base R
  - comparison
subtitle: ''
summary: ''
authors: []
lastmod: '2022-01-09T18:30:38+08:00'
featured: no
image:
  caption: ''
  focal_point: ''
  preview_only: no
projects: []
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

There are two functions in R that seems almost similar yet different: 

1. `fitted()`\
2. `predict()`

First let's prepare some data first.

```{r warning=FALSE, message=FALSE}
# Packages
library(dplyr)

# Data
set.seed(123)
dat <- 
  iris %>% 
  mutate(twoGp = sample(c("Gp1", "Gp2"), 150, replace = T), #create two group factor
         twoGp = as.factor(twoGp))
```

This is our data.
```{r}
summary(dat)
```


`fitted()` is used to get a predicted values or $\hat{y}$ based on the data. Let's see this on the logistic regression.

```{r}
logR <- glm(twoGp ~ ., family = binomial(), data = dat)
```

These are the fitted values.

```{r}
fitted(logR) %>% head()
```

For `predict()`, we have three types:

1. Response\
2. Link - default\
3. Terms

If no new data supplied to `predict()`, it will use the original data used to fit the model.

**1. Response**

The type response is identical to `fitted()`.

```{r}
predict(logR, type = "response") %>% head()
```

We can confirm this as below.

```{r}
all.equal(fitted(logR), predict(logR, type = "response"))
```

Thus, `fitted()` and `predict(type = "response")` give use predicted probabilities on the scale of the response variable. The first observation of this values can be interpreted as probability of Gp2 (Gp1 is a reference group) for first observation is 0.41. 

**2. Link**

`predict(type = "link")` gives us predicted probabilities on the logit scale or log odds prediction.

```{r}
predict(logR, type = "link") %>% head() #similar to predict(logR)
```

So, the log odds prediction of Gp2 for the first observation is -0.37. Hence, we can get the same values if we apply a [link function](https://en.wikipedia.org/wiki/Generalized_linear_model#Link_function) to the fitted values. 

The link function for logistic regression is:

$$
ln(\frac{\mu}{1 - \mu})
$$
So, we apply this link function to the fitted values.

```{r}
logOddsProb <- log(fitted(logR) / (1 - fitted(logR))) 
head(logOddsProb)
```

We can further confirm this as we did previously.

```{r}
all.equal(logOddsProb, predict(logR, type = "link"))
```

Also, we can conclude `predict(type = "link")` give use a fitted values before an application of link function (log odds).

**3. Terms**

Lastly, we have `predict(type = "terms")`. This type gives us a matrix of fitted values for each variable of each observation in the model on the scale of linear predictor. 

```{r}
predict(logR, type = "terms") %>% head() 
```

So, if we add up the values of the first observation and the constant (or intercept), we will get the same values as the log odds prediction (`predict(type = "link")`).

```{r}
predTerm <- predict(logR, type = "terms")
sum(predTerm[1, ], attr(predTerm, "constant")) #add up the first observation and the constant
logOddsProb[1]
```

Those values also similar to if we calculate manually using coefficient from the `summary()`.

$$
LogOdds(Gp2) = \beta_0 + \beta_1(Sepal.Length) + \beta_2(Sepal.Width) + 
$$
$$
\beta_3(Petal.Length) + \beta_4(Petal.Width) + \beta_5(Species)
$$
So, this is the values we get from the first observation.

```{r}
coef(logR)[1] + coef(logR)[2]*dat$Sepal.Length[1] + coef(logR)[3]*dat$Sepal.Width[1] + coef(logR)[4]*dat$Petal.Length[1] + coef(logR)[5]*dat$Petal.Width[1] + 0 #setosa species
```

However, in `predict(type = "terms")` the values is [centered](https://www.statology.org/center-data-in-r/), thus we have a different values for constant/intercept and for $\beta_1(Sepal.Length)$,$\beta_2(Sepal.Width)$ and so on. For example, the values for intercept for both models are:

```{r}
# Intercept/constant from predict(type = "terms")
attr(predTerm, "constant")
# Intercept/constant from summary()
coef(logR)[1]
```

References:

- <https://stackoverflow.com/a/12201502/11215767>\
- <https://stackoverflow.com/a/47854088/11215767>