---
title: Making maps with R (my first attempt ever!)
author: tengku-hanis
date: '2021-11-12'
slug: making-maps-with-r-my-first-attempt-ever
categories:
  - R
  - Epidemiology
  - Map
tags:
  - map
  - spatial
subtitle: ''
summary: ''
authors: []
lastmod: '2021-11-12T09:58:02+08:00'
featured: no
image:
  caption: ''
  focal_point: ''
  preview_only: no
projects: []
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

As written in the title of the post, this is my first try ever in making a map with R. I found a great data on the distributions of the clinic in Malaysia. The two types of clinic that we have here:

1) Klinik 1Malaysia (1Malaysia clinic)
2) Klinik Desa (Desa clinic)

Originally, these two data are a separated data. Both of the data can be downloaded from [here](https://www.data.gov.my/data/ms_MY/group/pemetaan). Also, I have uploaded the data into my [GitHub repo](https://github.com/tengku-hanis/clinic-data) for those interested. Klinik Desa data have a latitude and longitude information, but Klinik 1Malaysia does not.

These are the required packages.
```{r}
library(rworldmap) #to get a Malaysia map
library(tidyverse)
library(tidygeocoder) #to get latitude and logitude
```

Read the data.
```{r}
clinic1m <- read.csv("https://raw.githubusercontent.com/tengku-hanis/clinic-data/main/clinic1m.csv")
clinicDesa <- read.csv("https://raw.githubusercontent.com/tengku-hanis/clinic-data/main/clinicdesa.csv")
```

First, we need to get a latitude and longitude information for Klinik 1Malaysia data. So, we going to retrieve the coordinates based on the postal code, though this is not very accurate. We can use `tidygeocoder` for this.
```{r}
clinic1m2 <- 
  clinic1m %>%
  mutate(country = "malaysia") %>% 
  select(name, postcode, country) %>% 
  mutate(postcode = ifelse(nchar(postcode) == 4, paste0(0, postcode), postcode)) %>%
  geocode(postalcode = postcode, country = country, method = "osm")
```

Further checking on the data, we notice that 5 clinics have no coordinate info.
```{r}
clinic1m2 %>% filter(is.na(lat) | is.na(long))
```

## Some data pre-processing

So, I found this [data](https://www.listendata.com/2020/11/zip-code-to-latitude-and-longitude.html) after some googling time, which give coordinate based on the postal code. So, we going to add in the missing coordinate based on this online data.
```{r}
add_coord <- 
  read.table(header = T, text = "
postal_code    latitude   longitude
16070            6.0334    102.3499
26060            3.6228    102.3926
90700            5.8456    118.0571
26060            3.6228    102.3926")

clinic1m2 <- 
  clinic1m2 %>% 
  mutate(lat = ifelse(postcode %in% add_coord$postal_code, add_coord$latitude, lat), 
         long = ifelse(postcode %in% add_coord$postal_code, add_coord$longitude, long)) %>% 
  drop_na() #drop 2 clinic1m
```

Even after add in the missing coordinate, we still missing 2 coordinates. So, we going to drop those 2 clinics. Next, we combine both data.
```{r}
all_clinic <- 
  clinic1m2 %>% 
  mutate(Type = "1Malaysia") %>% 
  select(Type, lat, long) %>% 
  bind_rows(clinicDesa %>% 
              mutate(Type = "Desa", 
                     lat = latitude, 
                     long = longitude) %>% 
              select(Type, lat, long))
```

Let's try plotting the data first.
```{r}
ggplot(all_clinic, aes(long, lat, color = Type)) +
  geom_point() +
  theme_minimal() #should remove the isolated two data
```

We have 2 isolated points from Klinik Desa data. We will drop these 2 points as well.
```{r}
all_clinic2 <- all_clinic %>% filter(long > 25)
```

## Plotting the map

There are 2 ways to plot our data to Malaysia map, that we going to cover in this post.

### 1) map from `ggplot2` 

First, we need to get the map.
```{r}
global <- map_data("world") #get map
```

Once, we retrieved the map, we need to filter the region to Malaysia. The rest of the codes are `ggplot2` function as we know it.
```{r}
ggplot() + 
  geom_polygon(data = global %>% filter(region == "Malaysia"), aes(x=long, y = lat, group = group), 
               fill = "gray85") + 
  coord_fixed(1.3) +
  geom_point(data = all_clinic2, aes(x = long, y = lat, group = Type, color = Type, shape = Type)) +
  theme_void() + 
  xlab("Longitude") +
  ylab("Latitude") +
  labs(title = "Klinik 1Malaysia dan klinik desa di Malaysia", 
       subtitle = "(Data dikemaskini: Klnik 1Malaysia - 16 Mac 2021, Klnik desa - 9 Mac 2021)",
       caption = expression(paste(italic("Sumber data: https://www.data.gov.my/data/ms_MY/group/pemetaan"))), 
       color = "Jenis klinik:", 
       shape = "Jenis klinik:") +
  theme(plot.title = element_text(hjust = 0.5), 
        plot.subtitle = element_text(hjust = 0.5), 
        legend.position = "bottom") 
```

### 2) map from `rworldmap`

The flow is similar, we need to get the map first. Then, restrict the map to Malaysia region.
```{r}
world <- getMap(resolution = "low") #get map
msia <- world[world@data$ADMIN == "Malaysia", ]
```

The rest of the codes are similar to the first approach. But, we going to change the theme a bit.
```{r}
ggplot() +
  geom_polygon(data = msia, aes(x = long, y = lat, group = group), fill = NA, colour = "black") +
  geom_point(data = all_clinic2, aes(x = long, y = lat, group = Type, color = Type, shape = Type)) +
  coord_quickmap() + 
  theme_minimal() + 
  xlab("Longitude") +
  ylab("Latitude") +
  labs(title = "Klinik 1Malaysia dan klinik desa di Malaysia", 
       subtitle = "(Data dikemaskini: Klnik 1Malaysia - 16 Mac 2021, Klnik desa - 9 Mac 2021)",
       caption = expression(paste(italic("Sumber data: https://www.data.gov.my/data/ms_MY/group/pemetaan"))), 
       color = "Jenis klinik:", 
       shape = "Jenis klinik:") +
  theme(plot.title = element_text(hjust = 0.5), 
        plot.subtitle = element_text(hjust = 0.5), 
        legend.position = "bottom")
```

## Conclusion
The coordinates that we have are not as accurate as it should, or maybe there is something wrong that I miss along the way. As we can see, we have clinics on the ocean. As far as I know, we Malaysian are not that advanced yet. Also, noticed that we severely lacking clinics in Sarawak, given that our data is correct. 

