---
title: My first interactive map with {leaflet}
author: tengku-hanis
date: '2021-11-28'
slug: my-first-interactive-map-with-leaflet
categories:
  - R
  - Map
tags:
  - spatial
  - map
subtitle: ''
summary: ''
authors: []
lastmod: '2021-11-28T09:56:59+08:00'
featured: no
image:
  caption: ''
  focal_point: ''
  preview_only: no
projects: []
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

I have tried creating a map with ggplot2 [previously](https://tengkuhanis.netlify.app/post/making-maps-with-r-my-first-attempt-ever/). In this post, I will try to create an interactive map using `leaflet` package in R.

These are the required packages.

```{r}
library(tidyverse)
library(tidygeocoder)
library(leaflet)
```

So, I'm going to use a clinics location data in Malaysia. I already uploaded this data tomy [GitHub repo](https://github.com/tengku-hanis/clinic-data). I will skip the explanation for the pre-processing part, but it is the same pre-processing as my [previous post](https://tengkuhanis.netlify.app/post/making-maps-with-r-my-first-attempt-ever/).

```{r}
# Read the data
clinic1m <- read.csv("https://raw.githubusercontent.com/tengku-hanis/clinic-data/main/clinic1m.csv")
clinicDesa <- read.csv("https://raw.githubusercontent.com/tengku-hanis/clinic-data/main/clinicdesa.csv")
```

<details>
  <summary>Show code for pre-processing</summary>
```{r}
# Get the missing coordinate based on postal codes
clinic1m2 <- 
  clinic1m %>%
  mutate(country = "malaysia") %>% 
  select(name, postcode, country) %>% 
  mutate(postcode = ifelse(nchar(postcode) == 4, paste0(0, postcode), postcode)) %>%
  geocode(postalcode = postcode, country = country, method = "osm")

# Add coordinate from external sources for the still missing coordinates
add_coord <- 
  read.table(header = T, text = "
postal_code    latitude   longitude
16070            6.0334    102.3499
26060            3.6228    102.3926
90700            5.8456    118.0571
26060            3.6228    102.3926")

# Drop clinics with the still missing coordinate
clinic1m2 <- 
  clinic1m2 %>% 
  mutate(lat = ifelse(postcode %in% add_coord$postal_code, add_coord$latitude, lat), 
         long = ifelse(postcode %in% add_coord$postal_code, add_coord$longitude, long)) %>% 
  drop_na() #drop 2 clinic1m

# Bind the 2 data
all_clinic <- 
  clinic1m2 %>% 
  mutate(Type = "1Malaysia") %>% 
  select(name, Type, lat, long) %>% 
  bind_rows(clinicDesa %>% 
              mutate(Type = "Desa", 
                     lat = latitude, 
                     long = longitude) %>% 
              select(name, Type, lat, long)) %>% 
  mutate(name = str_to_title(name))
```
</details>

First, we going to plot the coordinates to see if there is anything strange.

```{r}
ggplot(all_clinic, aes(long, lat, color = Type)) +
  geom_point() +
  theme_minimal()
```

So, we are going to remove the two isolated points as seen from the plot.

```{r}
all_clinic2 <- all_clinic %>% filter(long > 25)
```

Once we have our data ready, we can supply to `leaflet`. We can choose the type of map from `addProviderTiles()`. Some need an api, but the one we choose here does not. We supply the longitude and latitude of our data to `addCircleMarkers()`, and `clusterOptions` to cluster our data.

```{r}
leaflet(all_clinic2) %>% 
  addProviderTiles(providers$Stamen.Watercolor) %>%
  addProviderTiles(providers$Stamen.TerrainLabels) %>%
  addCircleMarkers(~long, ~lat, 
                   clusterOptions = markerClusterOptions())
```

Next, we can add a label.

```{r}
labels <- 
  sprintf("<strong>%s</strong>", all_clinic$name) %>% 
  lapply(htmltools::HTML)
```

Also, we can add a minimap to our map. Here, I change the type of map to a more appropriate one.

```{r}
leaflet(all_clinic2) %>% 
  addProviderTiles(providers$OpenStreetMap) %>%
  addCircleMarkers(~long, ~lat, popup = ~labels, # popup add the label
                   clusterOptions = markerClusterOptions()) %>% 
    # add a minimap
  addMiniMap(tiles = providers$OpenStreetMap, zoomLevelOffset = -3)
```

Notice that the coordinates look more accurate as compared to the map I created with `ggplot2` previously.

References:

-   <https://lauriebaker.rbind.io/post/where_work/>
-   <https://laurielbaker.github.io/DSCA_leaflet_mapping_in_r/>
